### 2021-02-16 referee_response.md

> You report a final sample of 139 galaxies with 261 spiral arms, i.e. less than 2 arms per galaxy. Please give details of why this is so. Please also list how many of the galaxies are multiarmed, and state the number of arms.

When we looked into this we found this needed to be updated to 129 galaxies with 247 spiral arms which is closer to two arms per galaxy (1.9). We are unsure of the reason for this discrepancy, it may have been a typo, or left over from an earlier run.

Here is the complete breakdown of number of arms per galaxy:

```159 total galaxies, 129 of which with identified spiral arms
247 total spiral arms
238,433 total points
 30  galaxies had 0 arm(s)
 38  galaxies had 1 arm(s)
 68  galaxies had 2 arm(s)
 19  galaxies had 3 arm(s)
 4   galaxies had 4 arm(s)```

### 2021-02-16 referee_response.md

> You report a final sample of 139 galaxies with 261 spiral arms, i.e. less than 2 arms per galaxy. Please give details of why this is so. Please also list how many of the galaxies are multiarmed, and state the number of arms.

When we looked into this we found this needed to be updated to 129 galaxies with 247 spiral arms which is closer to two arms per galaxy (1.9). We are unsure of the reason for this discrepancy, it may have been a typo, or left over from an earlier run.

Here is the complete breakdown of number of arms per galaxy:

```159 total galaxies, 129 of which with identified spiral arms
247 total spiral arms
238,433 total points
 30  galaxies had 0 arm(s)
 38  galaxies had 1 arm(s)
 68  galaxies had 2 arm(s)
 19  galaxies had 3 arm(s)
 4   galaxies had 4 arm(s)

This has been entered into the text.```

### 2021-02-16 referee_response.md

> You report a final sample of 139 galaxies with 261 spiral arms, i.e. less than 2 arms per galaxy. Please give details of why this is so. Please also list how many of the galaxies are multiarmed, and state the number of arms.

When we looked into this we found this needed to be updated to 129 galaxies with 247 spiral arms which is closer to two arms per galaxy (1.9). We are unsure of the reason for this discrepancy, it may have been a typo, or left over from an earlier run.

Here is the complete breakdown of number of arms per galaxy:

159 total galaxies, 129 of which with identified spiral arms
247 total spiral arms
238,433 total points
 30  galaxies had 0 arm(s)
 38  galaxies had 1 arm(s)
 68  galaxies had 2 arm(s)
 19  galaxies had 3 arm(s)
 4   galaxies had 4 arm(s)

### 2021-02-16 referee_response.md

> Figure 8 has only one panel, so the sentences referring to the upper and lower panels of this figure should be rephrased.

This error has been corrected.

> The range of values of PHI_gal is from 8 to about 50 degrees, and there is no indication that the lower limit in the pitch angle is poorly determined. Note that Kennicutt (1981) find a lower limit of 0 degrees. Setting the lower limit to 15 degrees, so as just to be able to conclude that a uniform cot PHI_gal or cot PHI_arm distribution cannot be excluded is thus completely arbitrary. It would be best to display in a second panel the distribution of cot PHI_gal or cot PHI_arm with a lower limit of 8 degrees, so as to alert the reader to the strong influence of this assumption.

 The issue here is that the model itself includes an arbitrary lower pitch angle - Pringle & Dobbs (2019) make no attempt to give any physical motivation for what that minimum angle should be, and in their test against data chose 11.5 degrees as the limit - largely because that's where inspection of the distribution of cot(pitch angle) in the sample they investigate deviates from constant.

 We think the referee makes and excellent point that the previous explanation was a little confusing. What we have really done is to say this model fits the data IF the minimum pitch angle for winding spirals is around 15 degrees. It cannot be as low as 10 degrees. That's a specific observations that any more detailed model based on the simple ideas of Pringle & Dobbs (2019) would have to explain. We have made edits to the text to make this point.

### 2021-02-16 sections/abstract.tex
Edits about the minimum pitch angle in the cot-uniform model

% !TEX root = ../main.tex
Spiral structure is ubiquitous in the Universe, and the pitch angle of arms in spiral galaxies is an important observable in efforts to discriminate between different mechanisms of spiral arm formation and evolution. In this paper, we present a hierarchical Bayesian approach to galaxy pitch angle determination, using spiral arm data obtained through the \textit{Galaxy Builder} citizen science project. We present a new approach to deal with observed large variations in pitch angle between different arms in a single galaxy, which and obtains full posterior distributions on parameters. We make use of our fitted pitch angles to examine previously reported links between bulge and bar strength and pitch angle, finding no correlation in our data. We test a recent model for spiral arm winding, which predicts uniformity of the cotangent of pitch angle between some unknown upper and lower limits, with results supporting this model of transient and recurrent spiral pitch angle.

>>>>

% !TEX root = ../main.tex
Spiral structure is ubiquitous in the Universe, and the pitch angle of arms in spiral galaxies is an important observable in efforts to discriminate between different mechanisms of spiral arm formation and evolution. In this paper, we present a hierarchical Bayesian approach to galaxy pitch angle determination, using spiral arm data obtained through the \textit{Galaxy Builder} citizen science project. We present a new approach to deal with observed large variations in pitch angle between different arms in a single galaxy, which and obtains full posterior distributions on parameters. We make use of our fitted pitch angles to examine previously reported links between bulge and bar strength and pitch angle, finding no correlation in our data {\bf (with a caveat that we use observations proxies for both bulge size and bar strength which differ from other work)}. We test a recent model for spiral arm winding, which predicts uniformity of the cotangent of pitch angle between some unknown upper and lower limits, {\bf finding our observations are consistent with} this model of transient and recurrent spiral pitch angle {\bf as long as the pitch angle at which most winding spirals dissipate/disappear is larger than 10$^\circ$.}

### 2021-02-16 sections/results.tex
Edits about the minimum pitch angle in the cot-uniform model

In QSDW theory, the pattern speed $\Omega_p$ is a constant in R, as spiral arms obey rigid-body rotation. If $\Omega_p$ instead varies with radius we would expect $\cot{\phi}$ to be uniformly distributed between $\cot{\phi_\mathrm{max}}$ and $\cot{\phi_\mathrm{min}}$.

To test this theory, \citet{2019arXiv190910291P} used a Kolmogorov-Smirnov test to examine whether a sample of observed galaxy pitch angles was likely to have been drawn from a distribution uniform in its cotangent. Pitch angles were measured using discrete Fourier transformations in one- and two-dimensions, and as such do not account for inter-arm variations. They chose limits of $\cot{\phi} \in [1.00, 4.75]$ (roughly $11.9^\circ < \phi < 45.0^\circ$), motivated by examination of their data.

We perform a similar test in this work, using our sample and methods. We will make use of the marginalized Anderson-Darling test described above, and examine winding on a per-arm basis, as well as a per-galaxy basis. Observation of the distribution of arm pitch angles in our sample (Figure \ref{fig:pa-cot-distributions}) suggests limits of $15^\circ < \phi < 50.0^\circ$.

@@

This result suggests that we cannot rule out a cot-uniform source distribution for galaxy pitch angle, but the large uncertainty in $\phi_\mathrm{gal}$ makes it difficult to make any conclusive statements. This result is also highly sensitive to the lower limit of $\phi$: decreasing it to {10\degree} results in us rejecting the cot-uniform model at greater than the 0.1\% level for 96\% of the posterior samples. As we have no information available on the selection biases present for classification of extremely loose or tight spiral arms in \textit{Galaxy Builder}, we choose to keep the less strict limit of {15\degree}.

@@

\begin{figure*}
  \includegraphics[width=17.7cm]{plots/combined_cot_uniform_marginalized_tests.pdf}
  \caption{The results of a marginalized Anderson-Darling test for uniformity in $\cot$ for $\phi_\mathrm{gal}$ (blue) and $\phi_\mathrm{arm}$ (orange), with values corresponding to various confidence intervals shown. Moving rightwards on the x-axis implies greater confidence in rejecting the null hypothesis that the sample was drawn from a distribution uniform in $\cot$ between $15^\circ < \phi < 50.0^\circ$. In this instance, we would not be able to reject the null hypothesis at the 1\% level for either $\phi_\mathrm{gal}$ or $\phi_\mathrm{arm}$, meaning our sample is consistent with a cot uniform distribution.}
  \label{fig:ad-cot-test}
\end{figure*}

>>>>

In QSDW theory, the pattern speed $\Omega_p$ is a constant in R, as spiral arms obey rigid-body rotation. If $\Omega_p$ instead varies with radius we would expect $\cot{\phi}$ to be uniformly distributed between $\cot{\phi_\mathrm{max}}$ and $\cot{\phi_\mathrm{min}}$. {\bf The model presented in \citet{2019arXiv190910291P} does not give any physical justification for what $\cot{\phi_\mathrm{max}}$ and $\cot{\phi_\mathrm{min}}$ should be. }

To test this theory, \citet{2019arXiv190910291P} used a Kolmogorov-Smirnov test to examine whether {\bf a sample of 113} galaxies with measured pitch angles was likely to have been drawn from a distribution uniform in its cotangent. Pitch angles were measured {\bf by \citet{2019ApJ...871..194Y}} using discrete Fourier transformations in one- and two-dimensions, and as such do not account for inter-arm variations. They {\bf conclude the model works within limits of} $\cot{\phi} \in [1.00, 4.75]$ (roughly $11.9^\circ < \phi < 45.0^\circ$), motivated by examination of the data.

We perform a similar test in this work, using our sample and methods. We will make use of the marginalized Anderson-Darling test described above, and examine winding on a per-arm basis, as well as a per-galaxy basis. Observation of the distribution of arm pitch angles in our sample (Figure \ref{fig:pa-cot-distributions}) suggests {\bf they are close to uniform in cotangent} within limits of $15^\circ < \phi < 50.0^\circ$.

@@

This result suggests that we cannot rule out a cot-uniform source distribution for galaxy pitch angle, but the large uncertainty in $\phi_\mathrm{gal}$ makes it difficult to make any conclusive statements. This result is also highly sensitive to the lower limit of $\phi$: decreasing it to {10\degree} results in us rejecting the cot-uniform model at greater than the 0.1\% level for 96\% of the posterior samples. {\bf We can conclude from this that the \citet{2019arXiv190910291P} cot-uniform model is a good fit to the data, as long as the minimum pitch angle, $\cot{\phi_\mathrm{min}}$, at which the majority of winding dissipate and/or disappear is  $\phi_\mathrm{min} > $ {10\degree}, and more confidently $\phi_\mathrm{min} = $ {15\degree}.} \sout{As we have no information available on the selection biases present for classification of extremely loose or tight spiral arms in \textit{Galaxy Builder}, we choose to keep the less strict limit of {15\degree}.}

@@

\begin{figure*}
  \includegraphics[width=17.7cm]{plots/combined_cot_uniform_marginalized_tests.pdf}
  \caption{The results of a marginalized Anderson-Darling test for uniformity in $\cot$ for $\phi_\mathrm{gal}$ (blue) and $\phi_\mathrm{arm}$ (orange), with values corresponding to various confidence intervals shown. Moving rightwards on the x-axis implies greater confidence in rejecting the null hypothesis that the sample was drawn from a distribution uniform in $\cot$ between $15^\circ < \phi < 50.0^\circ$. In this instance, we would not be able to reject the null hypothesis at the 1\% level for either $\phi_\mathrm{gal}$ or $\phi_\mathrm{arm}$, meaning our sample is consistent with a cot uniform distribution. {\bf The larger error in $\phi_\mathrm{gal}$  means that this result is more significant for $\phi_\mathrm{arm}$, which is also physically motivated, as arms can wind independently.}}
  \label{fig:ad-cot-test}
\end{figure*}

### 2021-02-16 sections/summary.tex
Edits about the minimum pitch angle in the cot-uniform model

Our results are consistent with spiral winding of the form described by \citet{2019arXiv190910291P}, in which spiral arms are transient and recurrent, evolve through mechanisms such as swing-amplification \citep{1965MNRAS.130..125G} and which wind up over time. However, \textbf{this result depends on the range of pitch angles included in this test}. The assumptions of this model of spiral winding are also highly simplistic, and it leaves many unanswered questions: what determines the limits on $\phi$? Is the spiral arm equally apparent at all pitch angles, or is a selection effect present? This result is also not evidence against QSDW, as our distribution of pitch angles may be dictated by other factors such as disk shear.

>>>>

Our results are consistent with spiral winding of the form described by \citet{2019arXiv190910291P}, in which spiral arms are transient and recurrent, evolve through mechanisms such as swing-amplification \citep{1965MNRAS.130..125G} and which wind up over time, \textbf{as long as the minimum pitch angle before the majority of arms dissipate or disappear is $\phi_\mathrm{min} = $ {15\degree}}. The assumptions of this model of spiral winding are highly simplistic, and it leaves many unanswered questions: what determines the limits on $\phi$? Is the spiral arm equally apparent at all pitch angles, or is a selection effect present? \textbf{Our observations suggest that any further development of this model needs to predict that the minimum pitch angle, $\phi_\mathrm{min}>${10\degree}.} This result is also not evidence against QSDW, as our distribution of pitch angles may be dictated by other factors such as disk shear.
